get---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##### Set global options to save plots
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='figures/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

## Loading and preprocessing the data
```{r loadingdata}
#Set the Working directory to the right location
defaultWD <- "//Users/gamelord/Documents/OneDrive/Coursera.org/5. Reproducible research/RepData_PeerAssessment1"
setwd(defaultWD)
unzip("./activity.zip", files = NULL, list = FALSE, overwrite = TRUE,
      junkpaths = FALSE, exdir = "./", unzip = "internal",
      setTimes = FALSE)
#data without N.A
activity <- na.omit(read.csv(file="./activity.csv",header=TRUE))
```
## What is mean total number of steps taken per day?
```{r hist}
#Total number of steps taken per day
meanstep <- aggregate(activity$steps~activity$date,FUN=sum)
names(meanstep) <- c("date","steps")
mean1<-round(mean(meanstep[,2]))
median1<-median(meanstep[,2])
#Histogram of total number of steps taken each day
hist(meanstep[,2], main="Histogram of the total number of steps taken each day",xlab="Total steps each day")
abline(v=median(meanstep[,2]),col="red",lwd=1)
abline(v=mean(meanstep[,2]),col="blue",lwd=1)
```  
  
#####  Mean of the total number of steps taken per day `r mean1`  
#####  Median of the total number of steps taken per day `r median1`  

## What is the average daily activity pattern?
```{r timeseries}
avgstep <- aggregate(activity$steps~activity$interval,FUN=mean)
names(avgstep) <- c("interval","steps")
max1 <- avgstep[which.max(avgstep$steps),]
plot.ts(avgstep$steps,ylab="Average steps",main="Time series plot of 5-minute interval")
```
  
##### Which INTERVAL has the maximum number of steps on average `r max1[,1]`

## Imputing missing values
```{r}
#Full dataset
full <- read.csv(file="./activity.csv",header=TRUE)
summary(full$steps)
navalue <- full[which(is.na(full[,1])==TRUE),]
missingvalue <- nrow(navalue)
```
   
##### There're `r missingvalue` missing value(s) in the dataset
### Imputation medthod: I would like to fill the NA with average value of corresponding interval across the dataset
```{r imputation}
imputena <- merge(navalue,avgstep,by="interval")
imputena<-imputena[,-2]
names(imputena)<-c("internal","date","steps")
#Rearrange columns
imputena <- imputena[,c(3,2,1)]
#Round number the steps
imputena$steps<-round(imputena$steps)
#Convert num to int
imputena$steps <- as.integer(imputena$steps)
#Add back to non-NA dataset
names(imputena)<-names(activity)
filled <- rbind(activity,imputena)
#Total number of steps each day with filled data
totalsteps <- aggregate(filled$steps~filled$date,FUN=sum)
hist(totalsteps[,2], main="Histogram of total steps with imputation data",xlab="Total steps each day")
abline(v=median(totalsteps[,2]),col="red",lwd=1)
abline(v=mean(totalsteps[,2]),col="blue",lwd=1)
mean2 <- mean(totalsteps[,2])
median2 <- median(totalsteps[,2])
```
   
##### The new mean is `r mean2`  
##### The new median is `r median2`    
#### The new mean and median changed a little bit.The imputing missing data don't change the mean and median much

## Are there differences in activity patterns between weekdays and weekends?
```{r imputeWeekday}
#Transform date to proper date format
activity$date <- as.POSIXct(activity$date)
activity$weekdays <- weekdays(activity$date)
activity$weekend<-ifelse(activity$weekdays %in% c("Saturday","Sunday"),"Weekend","Weekday")
activity$id <- paste(activity$interval,activity$weekend)
#Average of steps each day by interval
avgweekday<- aggregate(activity$steps~activity$id,FUN=mean)
names(avgweekday)<-c("id","steps")
#Transform navalue dataset to impute missing steps count
navalue$date <- as.POSIXct(navalue$date)
navalue$weekdays <- weekdays(navalue$date)
navalue$weekend<-ifelse(navalue$weekdays %in%
c("Saturday","Sunday"),"Weekend","Weekday")
navalue$id <- paste(navalue$interval,navalue$weekend)
imputed<-merge(navalue,avgweekday,by="id")
imputed<-imputed[,-2]
imputed<-imputed[,c(6,2,3,4,5,1)]
imputed[,1]<-round(imputed[,1])
names(imputed)<-names(activity)
totalwd <- rbind(activity,imputed)
avgwd <- aggregate(totalwd$steps~totalwd$interval+totalwd$weekend,FUN=mean)
names(avgwd)<-c("Interval","Weekend","Steps")
#It took a while to see the Base plotting and GGplot can't produce exact look required. But Lattice system does it easily
```
```{r weekend}
library(lattice)
xyplot(Steps~Interval|Weekend,data=avgwd,type="l",layout=c(1,2),ylab="Number of steps",main="Steps between Weekend vs Weekdays")
```
   
### Weekdays seem to have a higher jumpstart but remain pretty lower compare to weekend. The number of steps in weekend seems to represent a higher physical activity during the day

